---
title: Filtering Spatial Data with Crosstalk
author: Matt Johnson
date: '2018-07-25'
tags:
  - crosstalk
  - leaflet
  - rstats
  - sf
slug: filtering-spatial-data
featured: "filter-header.png"
featuredalt: "wilderness"
featuredpath: "img"
categories:
  - spatial
type: post
---

In a [previous post](https://mrjoh3.github.io/2018/07/20/animating-spatial-data/) I wanted to look at the daily pattern of motor vehicle accidents across the state of Victoria. What I am really interested in though is the accidents in my local area. 

To do this I am going to use `leaflet`, `sf` to handle all the spatial data and `crosstalk` to filter the data visible in the map. The `crosstalk` package enables a level of interactivity that is usually reserved for more complex server (`shiny`) based applications.

The data is from the [VicRoads open data portal](https://vicroadsopendata-vicroadsmaps.opendata.arcgis.com/datasets/c2a69622ebad42e7baaa8167daa72127_0). The portal has an API but I have previously downloaded the entire data set as a shapefile. There are more than 75 thousand accidents recorded in the data set but I am only looking at accidents related to alcohol for the years 2015 through 2017.

```{r, warning=FALSE, message=FALSE}

library(sf)
library(dplyr)
library(leaflet)
library(crosstalk)
library(lubridate)
library(htmltools)


shp <- st_read('~/Documents/Matt/GIS/shp','Crashes_Last_Five_Years', quiet = TRUE) %>%
  mutate(SEVERITY = trimws(SEVERITY, 'both'),
         icons = as.factor(plyr::revalue(SEVERITY,
                              c('Serious injury accident' = 'serious',
                                'Fatal accident' = 'fatal',
                                'Other injury accident' = 'other'))),
         date = as.Date(as.character(ACCIDENT_D), '%d/%m/%Y'),
         year = year(date)) %>%
    filter(ALCOHOL_RE == 'Yes',
           year > 2014)

sd <- SharedData$new(shp)


```

Once the data is imported and cleaned it is inserted into a `crosstalk::SharedData$new()` object. This creates a common data input for all interactive objects such as UI elements and compatible widgets. For a complete list of compatible widgets see the [crosstalk reference page](https://rstudio.github.io/crosstalk/widgets.html).

Am minimal example can be just two lines of additional code which results in a UI slider connected to the date field and a leaflet map. Adjusting the slider will filter the markers visible in the map.

```r

filter_slider("date", "", sd, column=~date, step=10, width=800)
leaflet(sd) %>% addTiles() %>% addMarkers()


```

But let's add some additional layers of interactivity. `Crosstalk` has three UI filters built-in; slider, check-box and select. In the code below I am creating and saving the UI elements in order to have a little more control fitting them together.


```{r}

date_filter <- filter_slider("date", "", sd, column = ~date, 
                             step = NULL, width = '100%', dragRange = TRUE)

year_filter <- filter_checkbox("year", "Year", sd, group = ~year, 
                               inline = FALSE)

day_filter <- filter_checkbox("day", "Day of the Week", sd, group = ~DAY_OF_WEE, 
                              inline = FALSE)

severity_filter <- filter_checkbox("sev", "Severity", sd, group = ~SEVERITY, 
                                   inline = FALSE)

type_filter <- filter_checkbox("type", "Type", sd, group = ~ACCIDENT_1, 
                               inline = FALSE)


```

[Rstudio's leaflet](https://rstudio.github.io/leaflet/) package is the most ubiquitous method for incorporating an interactive map. In this post I am not going to discuss all the options I use but will include the code. To make the map look and feel how I want, I will use `addAwesomeMarkers`, `addPopups` and `addProviderTiles`.

Using custom marker colors and icons adds a lot of depth to a map and is essential if you have a large number of markers. But there are a couple of things to watch out for. 

First if you are matching icons to a column in your data it is best to create a new factor column to match. My initial column (`SEVERITY`) had spaces in the factor levels and this meant the icon labels would not match. 

Second, if you look at the code below you can see it is possible to combine multiple icon libraries. This does not make it a good idea. I found it difficult to determine which icons could be used without an account and which could not. This was particularly true for `glyphicons`; eventually I found [this link](https://getbootstrap.com/docs/3.3/components/).

Figuring out if I was having trouble with the factors or the icons was quite tedious, the intersection of these two issues made debugging difficult. Of course, with practice and patience I won't have this problem next time.

```{r}

iconSet = awesomeIconList(
  fatal = makeAwesomeIcon(markerColor = 'purple',
                          icon = 'ios-medical',
                          iconColor = 'black',
                          library = 'ion'),
  serious = makeAwesomeIcon(markerColor = 'darkred',
                            icon = 'ambulance', 
                            library = 'fa'),
  other = makeAwesomeIcon(markerColor = 'orange',
                          icon = 'plus', 
                          iconColor = 'black',
                          library = 'glyphicon')
)


```

Once you have the icon set you are ready to create a map. Now, I use `leaflet` maps a lot. They are incredibly versatile with a wealth of options (see [documentation](https://rstudio.github.io/leaflet/) to get started). They are also very simple. 

In the code below you can see my standard setup (excluding `addAwesomeMarkers()`). I save this as a snippet and insert it each time I make a map. This means that my maps are consistent and that I can get something done in a few minutes just focusing on the new layer.

```{r}

map <- leaflet(sd, width = '100%') %>% 
  addProviderTiles("CartoDB.Positron", group = 'Default') %>%
  addProviderTiles("Esri.WorldImagery", group = 'Aerial') %>%
  addProviderTiles("OpenStreetMap.Mapnik", group = 'Street') %>%
  addProviderTiles("OpenTopoMap", group = 'Terrain') %>%
  addScaleBar('bottomright') %>%
  addAwesomeMarkers(group = 'Accidents',
             icon = ~iconSet[icons],
             popup = ~paste0('<h3>Some Accident Details</h3><br>',
                             '<b>Type of Accident</b>: ', ACCIDENT_1, '<br>',
                             '<b>Light Conditions</b>: ', LIGHT_COND, '<br>',
                             '<b>Road Geometry</b>: ', ROAD_GEOME, "<br>",
                             '<b>Speed Zone</b>: ', SPEED_ZONE, '<br>',
                             '<b>Number of People</b>: ', TOTAL_PERS, '<br>',
                             '<b>Persons Injured or Fatality</b>: ', INJ_OR_FAT)
             ) %>%
  addLayersControl(
        baseGroups = c("Default", "Aerial", "Street", "Terrain"),
        overlayGroups = 'Accidents',
        options = layersControlOptions(collapsed = TRUE)
      )


```

Now we have all of the elements we need. These could have been added directly to the `rmarkdown` page with each code chunk, but wrapping them in some `htmltools` and `crosstalk` tags makes for a slightly cleaner user experience.

```{r, echo=TRUE}

tags$div(class="well well-lg",
         tagList(
           tags$h2('Alcohol Related Motor Vehicle Accidents 2014 - 2017'),
           date_filter,
           map,
           bscols(day_filter,
                  list(year_filter, severity_filter),
                  type_filter)
         )
)


```


