---
title: Filtering Spatial Data with Crosstalk
author: Matt Johnson
date: '2018-07-25'
tags:
  - crosstalk
  - leaflet
  - rstats
slug: filtering-spatial-data
type: post
---

In a [previous post](https://mrjoh3.github.io/2018/07/20/animating-spatial-data/) I wanted to look at the daily pattern of motor vehicle accidents across the state of Victoria. What I am really interested in though is the accidents in my local area. 

To do this I am going to use `leaflet`, `sf` to handle all the spatial data and `crosstalk` to filter and cut the data visible in the map. The `crosstalk` package enables a level of interactivity that is usually reserved for more complex server (`shiny`) based applications.

The data is from the [VicRoads open data portal](https://vicroadsopendata-vicroadsmaps.opendata.arcgis.com/datasets/c2a69622ebad42e7baaa8167daa72127_0). The portal has an API but I have preiously downloaded the entire dataset as a shapefile. There are more than 75 thousand accidents recorded in the dataset but I am only looking at accidents related to alcohol.

```{r, warning=FALSE, message=FALSE}

library(sf)
library(dplyr)
library(leaflet)
library(crosstalk)
library(leaflet.extras)
library(lubridate)


shp <- st_read('~/Documents/Matt/GIS/shp','Crashes_Last_Five_Years', quiet = TRUE) %>%
  filter(ALCOHOL_RE == 'Yes') %>%
  mutate(date = as.Date(as.character(ACCIDENT_D), '%d/%m/%Y'),
         year = year(date))

sd <- SharedData$new(shp)

```

Once the data is imported and cleaned it is inserted into a `crosstalk::SharedData$new()` object. This creates a common data input for all interactive objects such as UI elements and compatible widgets. For a complete list of compatible widgets see the [crosstalk reference page](https://rstudio.github.io/crosstalk/widgets.html).

Am minimal example can be just two lines of additional code which results in a UI slider conected to the date field and a leaflet map. Adjusting the slider will filter the markers visible in the map.

```r

filter_slider("date", "Date", sd, column=~date, step=10, width=800)
leaflet(sd) %>% addTiles() %>% addMarkers()


```

But let's add some additional layers of interactivity. `Crosstalk` has three UI filters built-in; slider, checkbox and select. In the code below I am creating and saving the UI elements in order to have a little more control fitting the together.


```{r}

date_filter <- filter_slider("date", "Date", sd, column = ~date, 
                             step = NULL, width = '100%', dragRange = TRUE)
year_filter <- filter_checkbox("year", "Year", sd, group = ~year, inline = FALSE)
day_filter <- filter_checkbox("day", "Day of the Week", sd, group = ~DAY_OF_WEE, inline = FALSE)
severity_filter <- filter_checkbox("sev", "Severity", sd, group = ~SEVERITY, inline = FALSE)
type_filter <- filter_checkbox("type", "Type", sd, group = ~ACCIDENT_1, inline = FALSE)

```

[Rstudio's leaflet](https://rstudio.github.io/leaflet/) package is the most ubiqitous method for incorporating an interactive map. In this post I am not going to discuss all the options I use but will include the code. But to make the map look and feel how I want, I will `addAwesomeMarkers`, `addPopups` and `addProviderTiles`.

```{r}

    iconSet = awesomeIconList(
      'primary school' = makeAwesomeIcon(markerColor = 'cadetblue',
                                         icon = 'graduation-cap', library = 'fa'),
      'secondary school' = makeAwesomeIcon(markerColor = 'darkred',
                                           icon = 'graduation-cap', library = 'fa'),
      'aged care' = makeAwesomeIcon(markerColor = 'red',
                                    icon = 'bed', library = 'fa'),
      'general hospital' = makeAwesomeIcon(markerColor = 'lightred',
                                           icon = 'h-square', library = 'fa'),
      'Other' = makeAwesomeIcon(markerColor = 'grey',
                                icon = 'question-circle', library = 'fa')
    )

map <- leaflet(sd, width = '100%') %>% 
  addProviderTiles("CartoDB.Positron", group = 'Default') %>%
  addProviderTiles("Esri.WorldImagery", group = 'Aerial') %>%
  addProviderTiles("OpenStreetMap.Mapnik", group = 'Street') %>%
  addProviderTiles("OpenTopoMap", group = 'Terrain') %>%
  addScaleBar('bottomright') %>%
  addAwesomeMarkers(group = 'Accidents',
             icon = ~iconSet[icon]) %>%
  addLayersControl(
        baseGroups = c("Default", "Aerial", "Street", "Terrain"),
        overlayGroups = 'Accidents',
        options = layersControlOptions(collapsed = TRUE)
      )
```



```{r, echo=FALSE}

library(htmltools)


tags$div(class="well well-lg",
         tagList(
           date_filter,
           map,
           bscols(day_filter,
                  list(year_filter, severity_filter),
                  type_filter)
         )
)



```


