---
title: Generate Web Link Preview Cards in Rmarkdown
author: Matt Johnson
date: '2022-02-02'
slug: generate-web-link-preview-cards-in-rmarkdown
categories:
  - web
  - R
tags:
  - rmarkdown
  - htmltools
  - webscape
  - metadata
description: |
  Have you ever noticed those fancy website previews that popup in some platforms like 
  twitter or slack when you include a link. How does the platform know exactly what 
  to display?
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

<link rel="stylesheet" href="card.css">




Recently I was creating a list of R resources and thought it might be interesting to replicate website previews seen when including links in twitter of slack posts. If I am honest though, a list of just URL's is neither interesting nor informative. And I did not want to spend all day looking up extra information about each resource.

Most modern websites include a large amount of metadata information embedded in the `<head></head>` tag. Where this metadata complies with the [Open Graph Protocol](https://ogp.me/) it is used to generate website previews and turns websites into data rich documents. For example if you right-click on this page select "inspect", you will find the following html `<meta>` tags:

```html

<meta property='og:title' content='Generate Web Link Preview Cards in Rmarkdown'>
<meta property='og:description' content='Have you ever noticed those fancy website previews that popup in some platforms like 
twitter or slack when you include a link. How does the platform know exactly what 
to display?
'>
<meta property='og:site_name' content='Restless Data'>
<meta property='og:type' content='article'>

```

The Open Graph tags are where the property attribute starts with "og:". There are also specific tags for twitter and facebook and we will read all of them. To use this metadata for our own previews within an rmarkdown document we will need to do a little web-scraping and generate some html. Lets start with the web-scraping.


## Extracting Metadata

The [rvest package](https://rvest.tidyverse.org/) makes it really easy to import the content of a webpage simply using `read_html()`. However, in this case we only need the metadata so there is no need to read the entire page. So I opted to use the `readLines()` function as it can limit the number of lines read in. Also, if you run `readLines()` within a `while` loop you can automatically stop reading once you find the end of the header. In general, this is not a particularly efficient way to read files, but as we mostly need to read less than 100 lines, it is fine.

Let's use the [Open Graph Protocol](https://ogp.me/) website as an example:

```{r read lines}

page <- file('https://ogp.me/', open = "r")

continue <- TRUE
doc <- as.character()

while(continue){

  l <- readLines(page, n = 1)
  continue <- !grepl('</head>', x = l, ignore.case = TRUE)
  doc <- c(doc, l)

}

doc
```

Now we can use `read_html()` to read the text and parse the tags; we want all of the meta tags plus the title tag. The inclusion of `paste0()` produces a character of length 1 which is required for `read_htlp()`, plus it ensures all the tags start and end points line up. The function `html_nodes()` is used to select only the nodes we are interested in. Some websites don't have a `meta` tag for the website title, so we select the `title` tag just in case.

```{r parse}

library(rvest)

nodes <- doc |>
  paste0(collapse = '') |>
  read_html() |>
  html_elements('meta, title')
nodes[1:7]

```
Now that we have all of the nodes, we can extract the parts of the tag with the information we need. The nodeset is a list so using the `purrr` package makes this extraction a lot easier. But you could also use `lapply` or `sapply`. As an end result we need to be able to lookup a metadata label and return the value. While this can easily be done in a `data.frame`, for lookups I find it a little clearer to convert to a list.

```{r}

library(purrr)
library(dplyr)

metadata <- map_dfr(c('property', 'name'), function(tag){
  nodes |>
    map_dfr(~ tibble(property = html_attr(.x, tag),
                     content = html_attr(.x, 'content'))) |>
    filter(!is.na(property))
})

metadata <- setNames(as.list(metadata$content), metadata$property)
```

Now that everything is working we can wrap it up into a function that takes any URL and returns a metadata lookup list. In the function below I am also pulling out the title tag. It can be used if the `og:title` is missing.


```{r metadata}

read_meta <- function(url){
  
  page <- file(url, open = "r")
  
  continue <- TRUE
  doc <- as.character()
  
  while(continue){
    l <- readLines(page, n = 1)
    continue <- !grepl('</head>', x = l, ignore.case = TRUE)
    doc <- c(doc, l)
  }

  close(page)
  
  nodes <- doc |>
    paste0(collapse = '') |>
    read_html() |>
    html_nodes('meta, title')
  
  metadata <- map_dfr(c('property', 'name'), function(tag){
    nodes |>
      map_dfr(~ tibble(property = html_attr(.x, tag),
                       content = html_attr(.x, 'content'))) |>
      filter(!is.na(property))
  })
  
  # check for a title tag
  title <- nodes |> html_nodes(xpath = "/html/head/title") |> html_text()
  title_tag <- if ('og:title' %in% metadata$property) 'title' else 'og:title'
  
  if (length(title) > 0) {
    metadata <- metadata |>
      add_row(property = title_tag,
              content = title)
  }
  
  metadata <- setNames(as.list(metadata$content), metadata$property)
  
  return(metadata)
  
}
```


## Render a Website Preview Card

Added a component like a card is much easier due to CSS frameworks like [Bootstrap](https://getbootstrap.com/docs/4.0/components/card/). In the card below I have a simple card created using [htmltools]() `tags` with predefined styles from Bootstrap.  

```{r}
library(htmltools)

card <- function(metadata){
  div(class = 'card text-white bg-secondary mb-3 border-light border-3',
            div(style = 'display: inline-grid; grid-template-columns: 55% 40%;padding-left:12px',
                div(class = 'card-body text-dark',
                    h2(class = 'card-title',
                       a(href = metadata$`og:url`,
                         metadata$`og:title`)
                       ),
                    p(class="card-text text-secondary",
                      metadata$`og:description`),
                    p(class="card-text text-secondary",
                      if (!is.null(metadata$author)) {
                        div("Author(s):", br(), metadata$author)
                      })
                  ),
                div(if (!is.null(metadata$`og:image`)) img(class="h-100", 
                                                           style = 'padding:20px;',
                                                           src = metadata$`og:image`, alt = metadata$`og:image:alt`) else '')
                )
            )
}
```


```{r card, echo=TRUE, message=FALSE, warning=FALSE}

card_url <- function(metadata){
  
  if (!is.null(metadata$`og:image:alt`)) {
    metadata$`og:image:alt` <- metadata$`og:title`
  }
  if (!is.null(metadata$author) & length(metadata$authur) > 1) {
    metadata$author <- metadata$author |> paste(collapse = '')
  }
  card(metadata)
}
```



```{r render a card}

read_meta('https://bookdown.org/yihui/rmarkdown/') |>
  card_url()

```










